---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![R CMD check](https://github.com/CmlMagneville/mFD/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CmlMagneville/mFD/actions/workflows/R-CMD-check.yaml)
[![Website deployment](https://github.com/CmlMagneville/mFD/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/CmlMagneville/mFD/actions/workflows/pkgdown.yaml)
[![License: GPL-2)](https://img.shields.io/badge/License-GPL%202-blue.svg)](https://choosealicense.com/licenses/gpl-2.0/)
[![Project Status: Active](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->


# SmaxN

The `SmaxN` package provides a unique function to compute maximisation of 
abundances index. It notably helps to compute the Synchronised maxN (SmaxN) based on
the use of temporally synchronised cameras.


## Installation


You can install the `SmaxN` package from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
## Install < remotes > package (if not already installed) ----
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

## Install < SmaxN > package from GitHub ----
remotes::install_github("CmlMagneville/SmaxN", build_vignettes = TRUE)
```


To compute the SmaxN and associated metrics, users need:

- a data frame gathering **abundances values through time** (rows) and cameras 
(columns) for the studied species
- a data frame gathering the **distance between camera pairs**
- the **maximal speed of the studied species** - established through laboratory
experiments, based on field performance or through video measurements if 
stereo cameras are used


## Citation

Please cite this package as:

_Magneville C., Brissaud C., Fleure V., Loiseau N., Claverie T., Villeger S._ 
(2024). A new framework for estimating abundance of animals using a network of cameras
_Limnology and Oceanography - Methods_
https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lom3.10606


## Contributions

SV, TC, NL, VF and CM developed the conceptual thinking of the SmaxN algorithm.
SV and CM developed the package architecture and CM coded the functions. CB, SV and
CM tested the functions. SV and CM contributed to writing of helps and tutorials.
