---
title: "How to compute the SmaxN metric?"
author: "Camille Magneville"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to compute the SmaxN metric?}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


The goal of this vignette is to quickly explain the goals of the `SmaxN` metric 
and how to simply compute it using the `SmaxN` package.

<br>


# What is the SmaxN metric about? 


The SmaxN is a metric which uses a network of synchronised cameras to estimate  
species abundance using the distances between the cameras and species maximal speed.
It relies on the simple principle that <span style="color: lightseagreen;">a mobile 
individual can not be in two places at the same time</span>.

<br>

The SmaxN framework needs only few conditions:

* <span style="color: lightseagreen;">a network of temporally synchronised cameras</span>: Surveys should be based on a set of at least two fixed cameras
which **field of view do not overlap**. The temporal synchronisation should be 
realised at a coherent precision according to the study design (at which 
precision individuals are counted and what distance separate the cameras). 
Basically, you just need to be able to link the frames from one camera to the 
frames of the other cameras. The distances between each pair of cameras must
be known and you must build a **distance data frame** with cameras in rows and 
columns.

* <span style="color: lightseagreen;">abundance through time</span>: abundance 
of the studied species should be retrieved through time. Thus for the studied 
species, there is a matrix with recorded abundance for each time step (in rows) 
on each camera (in columns) called the **camera × time abundance data frame**.

* <span style="color: lightseagreen;">species maximal speed</span>: the species 
maximal speed should be retrieved through experiments or using data available.
Yet to our knowledge, speed data is only available for a small number of species
[Husak et al. 2006](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.0014-3820.2006.tb01849.x) (collared lizard Crotaphytus collaris), [Alexander et al. 1977](https://zslpublications.onlinelibrary.wiley.com/doi/10.1111/j.1469-7998.1977.tb04188.x) (ten species of african ungulates), [Layne and Benton 1954](https://academic.oup.com/jmammal/article-abstract/35/1/103/847724?redirectedFrom=fulltext) (fourteen species of small mammals), [Hirt et al. 2017](https://pubmed.ncbi.nlm.nih.gov/29046579/) (474 terrestrial species), [Fulton 2007](https://link.springer.com/article/10.1007/s00338-007-0195-0) (117 coral reef fish species belonging to 10 families). If data is missing, we recommend to 
use **conservative estimates** that is the maximum of maximal speed among 
species from the same clade (e.g. family or order)

<br>

**NB**: The species maximal speed and the distances between the cameras must be
given using the same metric units (*i.e.* if using meters for the distance,
then use meters per seconds (for instance) for the speed)

<br>

Using the minimal time it takes for an individual of the studied species to go 
from one camera to another camera (using species maximal speed and distance 
between the cameras), the `SmaxN` algorithm find the best combination of 
abundances seen across time and cameras to get the highest possible estimate
of abundance.

<br>


# Let's compute the SmaxN & other abundance estimates


<br>

We will here work with a simple example.

First, let's build the distance data frame and the abundance data:


```{r}
# Build distance dataframe for the example:
dist_df_ex <- data.frame("A" = c(0, 2, 5, 5), "B" = c(2, 0, 5, 5), 
                         "C" = c(5, 5, 0, 4), "D" = c(5, 5, 4, 0))
rownames(dist_df_ex) <- c("A", "B", "C", "D")

# Build distance dataframe for the example:
abund_df_ex <- data.frame("A" = c(0, 1, 3, 7, 2, 2, 3, 0, 6, 2, 0, 1), 
                          "B" = c(2, 2, 2, 2, 0, 0, 0, 0, 1, 2, 1, 0), 
                          "C" = c(2, 0, 1, 0, 0, 4, 2, 2, 3, 0, 0, 4), 
                          "D" = c(0, 1, 0, 1, 0, 6, 1, 1, 6, 4, 2, 1))
dist_df_ex
```

In the distance dataframe there are for instance 2m from camera A to 
camera B, 5m from camera A to camera C and D *etc.*

<br>

```{r}
abund_df_ex
```

For instance, no individual is seen on the camera A on the first timestep, one individual is seen on the camera A on the second timestep, three individuals are 
seen on the camera A on the third time step *etc.*

<br> 

We can now call the package and apply the `SmaxN.computation` function to compute the `SmaxN` metric
among other metrics:


```{r}

SmaxN_results <- SmaxN::SmaxN.computation(abund_df = abund_df_ex, 
                                          speed = 1.6, 
                                          dist_df = dist_df_ex)
```

<br>

Let's now have a look at the results:

* the `maxN` value which is the highest abundance seen amon all cameras and 
time steps:

```{r}
SmaxN_results$maxN
```

* the `SmaxN` value: we can see here that by using the `SmaxN` metric, the 
abundance estmation is more than doubled than when using the `maxN` metric

```{r}
SmaxN_results$SmaxN
```

* the values of abundances along cameras and time steps which have led to the
`SmaxN` metric:

```{r}
SmaxN_results$path_saved
```

* the number of combinaisons of time steps and cameras which have led to the 
`SmaxN` value:

```{r}
SmaxN_results$number_SmaxN_path
```

* the `SmaxN_timestep` which is the maximal value of the sum of abundances 
across all cameras for each time step: Here this metrix equals 16 which is obtained through the ninth time step (6 + 1 + 3 + 6):

```{r}
SmaxN_results$SmaxN_timestep
```

* the maximal values obtained for each camera across all time steps:

```{r}
SmaxN_results$maxN_cam
```

* the maximal values obtained for each time step across all cameras:

```{r}
SmaxN_results$maxN_timestep
```


<br />

For further explanations, a paper see the paper associated with this 
framework: https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lom3.10606


<br>

# References

<br />

- Alexander _et al._ (1977) 
Fast locomotion of some African ungulates
_Journal of Zoology_, **183**, 291-300.

- Fulton (2007) 
Swimming speed performance in coral reef fishes: field validations reveal 
distinct functional groups
_Coral Reefs_, **26**, 217-228.

- Hirt _et al._ (2017) 
A general scaling law reveals why the largest animals are not the fastest 
_Nat. Ecol. Evol._ **1**, 1116–1122.

- Husak _et al._ (2006) 
Faster lizards sire more offspring: sexual selection on whole-animal performance 
_Evolution_ **60**, 2122–2130.

- Layne & Benton (1954) 
Some Speeds of Small Mammals
_Journal of Mammology_ **1**, 103–104.
